\renewcommand{\thefootnote}{\roman{footnote}}
\pagestyle{headings}
\markright{The Gospel of St. John - Patristic Commentary}
%\newcommand{\vn}[1]{\textsuperscript{#1}\hspace{0.15em}\ignorespaces}
%\newcommand{\verseNumber}[1]{\textsuperscript{#1}\hspace{0.2em}}

\newcommand{\vn}[1]{\textsuperscript{\textcolor{red}{\textbf{#1}}}\hspace{0.15em}\ignorespaces}
%\newcommand{\vn}[1]{\textsuperscript{\textbf{#1}}\hspace{0.15em}\ignorespaces}
j\marginpar{This is a text hehehe jlkdfjdslkj lkdsfjdslk }
\begin{multicols}{2}
\section*{JOHN}
    \lettrine[lines=3,slope=4pt,nindent=-3pt]{I}{n} the beginning\footnotemark[1] was the Word\footnotemark[2], and the Word was with\footnotemark[3] God, and the Word was\footnotemark[4] God.\vn{2}He was in the beginning with God.\vn{3}All things\footnotemark[5] were made through Him, and without Him nothing was made that was made.\vn{4}In Him was life, and the life was the light of men.\vn{5}And the light shines in the darkness, and the darkness did not comprehend it.
\vfill\null
\columnbreak

\section*{ΙΩΑΝΝΗΝ}
	\lettrine[lines=3,slope=4pt,nindent=-3pt]{Ἐ}{ν} ἀρχῇ\footnotemark[1] ἦν ὁ λόγος\footnotemark[2], καὶ ὁ λόγος ἦν πρὸς\footnotemark[3] τὸν θεόν, καὶ θεὸς ἦν\footnotemark[4] ὁ λόγος.\vn{2}οὗτος ἦν ἐν ἀρχῇ πρὸς τὸν θεόν.\vn{3}πάντα\footnotemark[5] δι’ αὐτοῦ ἐγένετο, καὶ χωρὶς αὐτοῦ ἐγένετο οὐδὲ ἕν. ὃ γέγονεν \vn{4}ἐν αὐτῷ ζωὴ ἦν, καὶ ἡ ζωὴ ἦν τὸ φῶς τῶν ἀνθρώπων·\vn{5}καὶ τὸ φῶς ἐν τῇ σκοτίᾳ φαίνει, καὶ ἡ σκοτία αὐτὸ οὐ κατέλαβεν.
\end{multicols}

\footnotetext[1]{Beginning - (Gr. \textit{αρχη})\\
\textit{\textbf{St. Basil the Great, Hexameron}} - ``The beginning, in effect, is indivisible and instantaneous. The beginning of the road is not yet the road, and that of the house is not yet the house; so the beginning of time is not yet time and not even the least particle of it. If some objector tell us that the beginning is a time, he ought then, as he knows well, to submit it to the division of time — a beginning, a middle and an end. Now it is ridiculous to imagine a beginning of a beginning. Further, if we divide the beginning into two, we make two instead of one, or rather make several, we really make an infinity, for all that which is divided is divisible to the infinite...''\\
\textit{\textbf{St. John Chrysostom, Homily 2}} - ``For the intellect, having ascended to \textit{``the beginning''}, enquires what \textit{``beginning''}; and then finding the \textit{``was''} always outstripping its imagination, has no point at which to stay its thought; but looking intently onwards, and being unable to cease at any point, it becomes wearied out, and turns back to things below. For this \textit{``was in the beginning,''} is nothing else than expressive of ever being and being infinitely.''\\
\textit{\textbf{St. Cyril of Alexandria, Homily 2}} - ``
Than the beginning is there nothing older, if it have, retained to itself, the definition of the beginning (for a beginning of beginning there cannot be); or it will wholly depart from being in truth a beginning, if something else be imagined before it and arise before it. Otherwise, if anything can precede what is truly beginning, our language respecting it will go off to infinity, another beginning ever cropping up before, and making second the one under investigation.
There will then be no beginning of beginning, according to exact and true reasoning, but the account of it will recede unto the long-extended and incomprehensive. And 12 since its ever-backward flight has no terminus, and reaches up to the limit of the ages, the Son will be found to have been not made in time, but rather invisibly existing with the Father: for in the beginning was He. But if He was in the beginning, what mind, tell me, can over-leap the force of the was? When will the was stay as at its terminus, seeing that it ever runs before the pursuing reasoning, and springs forward before the conception that follows it?
}


%\footnotetext[2]{Word - (Gr. \textit{λογος})\\
%\textit{\textbf{St. Irenaeus, Against Heresies (Book II, Chapter 2)}} -
%``And this is He of whom the Scripture says, And God formed man, taking clay of the earth, and breathed into his face the breath of life. Genesis 2:7 It was not angels, therefore, who made us, nor who formed us, neither had angels power to make an image of God, nor any one else, except the Word of the Lord, nor any Power remotely distant from the Father of all things.''
%\\
%\textit{\textbf{St. Augustine of Hippo, Tractates on the Gospel of John}} - ``This Word is unchangeable; although changeable things are made by it, the Word itself is unchangeable''
%\\
%\textit{\textbf{St. Justin Martyr, The First Apology}} -
%``Himself, who took shape, and became man, and was called Jesus Christ''}
%\footnotetext[3]{With - (Gr. \textit{πρὸς})}
%\footnotetext[4]{Was - (Gr. \textit{ἦν}) The Word is co-equal and co-eternal with the Father - sharing the same divine nature.}
%\footnotetext[5]{All things - (Gr. \textit{πάντα})\\
%\textit{\textbf{St. Irenaeus, Against Heresies (Book II, Chapter 2)}} - ``Now, among the all things our world must be embraced. It too, therefore, was made by His Word, as Scripture tells us in the book of Genesis that He made all things connected with our world by His Word. David also expresses the same truth [when he says] (Psalm 33:9) For He spoke, and they were made; He commanded, and they were created.''}
%For if God Is, as certainly He Is, then nothing was before Him. If He is Creator of all things, He must be first; if Master and Lord of all, then all, both creatures and ages, are after Him.


%Now, among the all things our world must be embraced. It too, therefore, was made by His Word, as Scripture tells us in the book of Genesis that He made all things connected with our world by His Word. David also expresses the same truth [when he says] For He spoke, and they were made; He commanded, and they were created.
